---
title: "SQL"
output: html_document
---

##Filtering Columns and Rows

With SQL you can filter columns and rows by using SELECT and WHERE. Let's look at an example using the Lahman database. So we should first load that database, along with the package sqdlf.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

####Example 1 
Suppose now I would like to see the home run totals for the 1927 Yankees. I could write the following query:

```{r}
query<-"select playerID,yearID,teamID,HR from Batting
where teamID='NYA'and yearID=1927"
sqldf(query)
```

####Example 2
We want to find instances where the Yankees have hit 40 or more home runs. I would write the query this way:
```{r}
query<-"Select playerID,yearID,teamID,HR from Batting
Where teamID='NYA' and HR>=40"
sqldf(query)
```

####Example 3
Now we want to see instances when a player had more than 40 home runs but less than 60 strikeouts. This is how the query would look:
```{r}
query<-"select playerID,yearID,teamID,HR,SO from Batting
where HR>=40 and SO<60"
sqldf(query)
```

####Example 4
Let's do a tougher example. We want to find instances where the Phillies in the 1970's hit more than 30 home runs. The tricky part is figuring out how to display data between 1970 and 1979. This is how I wrote the query:
```{r}
query<-"Select playerID,teamID,yearID,HR from Batting 
where teamID='PHI' and yearID<1980 and yearID>1969 and HR>30"
sqldf(query)
```

##Order by 

With SQL, you can also order the rows of your output.

####Example 1
Suppose we would like to see all instances of a player hitting more than 50 home runs with the highest home run totals at the top. The query is written like this:
```{r}
query<-"select playerID,teamID,yearID,HR from Batting
Where HR>50
Order by HR DESC"
sqldf(query)
```

####Example 2
For our next example, find instances of a player striking out less than 10 times. Make sure each player has had at least 400 at bats (AB). Order by having the least number of strikeouts at the top. Here's how the query should look:
```{r}
query<-"select playerID,teamID,SO,AB from Batting
Where AB>=400 and SO<10
Order by SO"
sqldf(query)
```

##Aggregation

This section is about grouping rows and columns together.

####Example 1
Our first example asks us to find all of Babe Ruth's home runs. If we write the query like this:
```{r}
query<-"select playerID,HR from Batting
where playerID='ruthba01'"
sqldf(query)
```
There are multiple rows with the same row title. To group all of these totals together, we have to use "group by". We also have to add all the home runs together by typing "sum" next to HR like so:
```{r}
query<-"select playerID,sum(HR) from Batting
where playerID='ruthba01'
group by playerID"

sqldf(query)
```

####Example 2
Find the career home run totals for all players but limit the output to total 600 or more. Order by career total. When you group before you filter you use "having" instead of "where."
```{r}
query<-"select playerID,sum(HR) from Batting
group by playerID
having sum(HR)>=600
order by sum(HR) DESC"

sqldf(query)
```

####Example 3
Now, find the players who have averaged more than 30 home runs per year throughout their career. Let's order by having the highest average at the top. 
```{r}
query<- "select playerID,avg(HR) from Batting
group by playerID
having avg(HR)>30
order by avg(HR) DESC"

sqldf(query)
```

##Inner Join

All this time we have been workign with one table, but you are also able to join two different tables together to find new information.

####Example 1
Instead of just playerID, we want a first and last name. We also want teamID, yearID, and HR. The player should be Babe Ruth. The table called Master has first and last name information. We use "inner join master" to join the Batting table and the Master table together. Then we use "on" to specify which variable joins the tables. So the quuery should look like this:
```{r}
query<-"select nameFirst,nameLast,teamID,yearID,HR
from Batting inner join Master
On Batting.playerID=Master.playerID
where Batting.playerID='ruthba01'"

sqldf(query)
```

####Example 2
Find all instances of players hitting more than 50 home runs. only this time, include first and last names, team, year, and home runs. If you want you can order them too. We use Batting._______  or Master._______to specify which table the column is coming from.
```{r}
query<-"select teamID,yearID,HR,nameFirst,nameLast
from Batting inner join Master
on Batting.playerID=Master.playerID
where Batting.HR>50
order by HR"

sqldf(query)
```

####Example 3
This is a short but tough example. We are using a different table called Teams this time. We want Babe Ruth with playerID, team name, Year, and HR. The tricky part is that two columns (teamID and YearID) in both tables are the same, not just one. So we have to specify that.
```{r}
query<-"select playerID,name,Batting.yearID,Batting.HR
from Batting inner join Teams
on Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
where playerID='ruthba01'"

sqldf(query)
```

